# Results {.unnumbered}

```{r, message=FALSE}
#| echo: FALSE
spotify_songs <- readr::read_csv("Data/scored_spotify.csv", show_col_types = FALSE)
```

```{r, message=FALSE}
#| label: fig-plots
#| fig-cap: "Duration and Intensity by Popularity Level" 
#| fig-subcap:
#|   - "Duration"
#|   - "Intensity" 
#| layout-ncol: 2
#| echo: FALSE

# load ggplot2
library(ggplot2)
spotify_songs$popularity_factor = factor(spotify_songs$popularity)

p <- ggplot(spotify_songs, aes(x=popularity_factor, y=duration_mins)) + geom_violin(trim=FALSE) + labs(title="Plot of duration by popularity level",x="Popularity level", y = "Duration in minutes")
p + geom_boxplot(width=0.1)

q <- ggplot(spotify_songs, aes(x=popularity_factor, y=intensity)) + geom_violin(trim=FALSE) + labs(title="Plot of intensity by popularity level",x="Popularity level", y = "Intensity")
q + geom_boxplot(width=0.1)
```

```{r, message=FALSE}
#| echo: FALSE
minutes_popularity = cor.test(spotify_songs$duration_mins,spotify_songs$popularity,method="spearman",exact=FALSE)

intensity_popularity = cor.test(spotify_songs$intensity, spotify_songs$popularity, method="spearman", exact=FALSE)
```

From @fig-plots we see that the distributions of intensity and duration are visibly skewed within popularity classes. As a result, we used a Spearman rank correlation to test each hypothesis. In @tbl-cor we see that both p-values are well below our pre-established cutoff of 0.05. Since we are doing multiple statistical tests, we could adjust the p-values to account for this. However, since there are only two tests and the p-values are several orders of magnitude smaller than our cutoff, this is unnecessary. Both duration and intensity were negatively correlated with popularity to a weak degree; neither coefficient was on the expected scale. Since intensity was a composite measure, exploration of how the components relate to popularity could explain why the direction of correlation was unexpected.

| Factor    | p-value                          | Coefficient                       |
|:---------------|:---------------------------|:----------------------------|
| Duration  | `r minutes_popularity$p.value`   | `r minutes_popularity$estimate`   |
| Intensity | `r intensity_popularity$p.value` | `r intensity_popularity$estimate` |

: Spearman Rank Correlation Results {#tbl-cor}
