# Results {.unnumbered}

```{r, message=FALSE}
#| echo: FALSE
spotify_songs <- readr::read_csv("Data/scored_spotify.csv", show_col_types = FALSE)
```

```{r, message=FALSE}
#| label: fig-plots
#| fig-cap: "Duration and Intensity by Popularity Level" 
#| fig-subcap:
#|   - "Duration"
#|   - "Intensity" 
#| layout-ncol: 2
#| echo: FALSE

# load ggplot2
library(ggplot2)
spotify_songs$popularity_factor = factor(spotify_songs$popularity)

p <- ggplot(spotify_songs, aes(x=popularity_factor, y=duration_mins)) + geom_violin(trim=FALSE) + labs(title="Plot of duration by popularity level",x="Popularity level", y = "Duration in minutes")
p + geom_boxplot(width=0.1)

q <- ggplot(spotify_songs, aes(x=popularity_factor, y=intensity)) + geom_violin(trim=FALSE) + labs(title="Plot of intensity by popularity level",x="Popularity level", y = "Intensity")
q + geom_boxplot(width=0.1)
```

```{r, message=FALSE}
#| echo: FALSE
minutes_popularity = cor.test(spotify_songs$duration_mins,spotify_songs$popularity,method="spearman",exact=FALSE)

intensity_popularity = cor.test(spotify_songs$intensity, spotify_songs$popularity, method="spearman", exact=FALSE)
```

From @fig-plots we see that the distributions of intensity and duration are visibly skewed within popularity classes. As a result, we used a Spearman rank correlation to test each hypothesis. In @tbl-cor we see that both p-values are well below our pre-established cutoff of 0.05. Since we are doing multiple statistical tests, we could adjust the p-values to account for this. However, since there are only two tests and the p-values are several orders of magnitude smaller than our cutoff, this is unnecessary and would not change the conclusions drawn from this analysis. Both duration and intensity were weakly negatively correlated with popularity; neither coefficient was on the expected scale. Since intensity was a composite measure, exploration of how the individual components, including energy, tempo, and loudness, relate to popularity could explain why the direction of correlation was unexpected.

| Factor    | p-value                                              | Coefficient                                         |
|:-----------------|:--------------------------|:--------------------------|
| Duration  | `r signif(minutes_popularity$p.value, digits = 3)`   | `r signif(minutes_popularity$estimate, digits=3)`   |
| Intensity | `r signif(intensity_popularity$p.value, digits = 3)` | `r signif(intensity_popularity$estimate, digits=3)` |

: Spearman Rank Correlation Results {#tbl-cor}
